---
title: 'STAT 447: Exercise 6'
author: 'Kevin Liu (94200474)'
date: "16 February, 2025"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#if you do not have the package, type install.packages("name_of_the_package")
library(knitr)
library(tseries)
```

### Question 1

$\mu_e\sim$ Unif$(0,1)$
    
$\lambda_e\sim$ Exp$(0.01)$

$\mu_p\sim$ BetaMP$(0.1,15)$

$\lambda_p\sim$ Exp$(0.01)$

$e_{Pfizer},e_{Moderna}|\mu_e,\lambda_e\sim$ BetaMP$(\mu_e,\lambda_e)$

$p_{Pfizer},p_{Moderna}|\mu_e,\lambda_e\sim$ BetaMP$(\mu_p,\lambda_p)$

$n_c|p\sim$ Binom$(t_c,p)$

$n_v|e,p\sim$ Binom$(t_v,p(1-e))$

### Question 2

```{r, echo=TRUE}
# this is where your R code goes
source("../../simple.R")

BetaMP = function(mean, precision){
  Beta(mean*precision, (1-mean)*precision)
}

model = function() {
  mu_e = simulate(Unif(0,1))
  lambda_e = simulate(Exp(0.01))
  
  mu_p = simulate(BetaMP(0.1, 15))
  lambda_p = simulate(Exp(0.01))
  
  eff = list()
  inc = list()
  
  eff[["Pfizer-BioNTech"]] = simulate(BetaMP(mu_e, lambda_e))
  inc[["Pfizer-BioNTech"]] = simulate(BetaMP(mu_p, lambda_p))
  
  eff[["Moderna-NIH"]] = simulate(BetaMP(mu_e, lambda_e))
  inc[["Moderna-NIH"]] = simulate(BetaMP(mu_p, lambda_p))
  
  for (i in seq_along(vaccines)) {
    vaccine = vaccines$trials[i]
    arm = vaccines$arms[i]
    size = vaccines$groupSizes[i]
    cases = vaccines$numbersOfCases[i]
    
    if (arm == "control") {
      observe(cases, Binom(size, inc[[vaccine]]))
    } else {
      observe(cases, Binom(size, inc[[vaccine]] * (1 - eff[[vaccine]])))
    }
  }

  return(eff[["Moderna-NIH"]] > eff[["Pfizer-BioNTech"]])
}
```

### Question 3

```{r, echo=TRUE}
# this is where your R code goes
vaccines = read.csv("vaccines.csv")
vaccines$groupSizes = as.double(vaccines$groupSizes) # needed due to bug in Binom code
```

### Question 4

```{r, echo=TRUE}
# this is where your R code goes
set.seed(2)
prob = posterior(model, 10000)
print(prob)
```

I am not sure if it is reliable because for the code above, when I set the seed to 1, it would fail for some reason. If this wasn't the case, I would say SNIS is reliable.