---
title: 'STAT 443: Assignment 2'
author: 'Kevin Liu (94200474)'
date: "March 14, 2025"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#if you do not have the package, type install.packages("name_of_the_package")
library(knitr)
```

For this project, as suggested on Piazza, I will look at Bayesian classification methods on the crime category. Alternatively, if that does not work out, I will use time series to find potential trends/seasonal effects on crime statistics (potentially splitting by neighborhood). Data is from [here](https://geodash.vpd.ca/opendata/#)

```{r, echo=TRUE}
crime_data = read.csv("crime_data/crime_data.csv")
head(crime_data)
```

A second data set that I could potentially use is from the WHO, with road traffic accidents for people aged 15-24. I could also use time series on this data set, or potentially apply a simple model and improved model. Data is from [here](https://platform.who.int/mortality/themes/theme-details/topics/indicator-groups/indicator-group-details/MDB/road-traffic-accidents)

```{r, echo=TRUE}
headers = read.csv("WHO.csv", skip = 6, header = F, nrows = 1, as.is = T)
WHO_data = read.csv("WHO.csv", skip = 7, header = F)
colnames(WHO_data)= headers
head(WHO_data)
```